---
title: "Data cleanup with RStudio"
author: "Nesserine ELMEDIOUNI"
date: "13/09/2020"
output: html_document
---

```{r setup, include=FALSE}
```
# Task 1: Read in unclean data and view from a high level view --#
```{r setup, include=FALSE}
setwd("")
getwd()

AutoSett <- read.csv("AutoSettlements_Messy.csv") 
AutoSett
AutoSett <- read.csv("AutoSettlements_Messy.csv", na.strings=c("")) #Replace any empty cells with NA
AutoSett
```
### Look at data at a high level
### How many rows?
### How many columns?
### What is the data type for each column?
```{r setup, include=FALSE}
head(AutoSett)
tail(AutoSett)
nrow(AutoSett)
ncol(AutoSett)
str(AutoSett)
summary(AutoSett)
```
# Task 2: Filter missing data --#
```{r setup, include=FALSE}
```
###Look for 60+ year olds, males with a settlement
###Hint: Use which() function to handle NAs
```{r setup, include=FALSE}
AutoSett[(AutoSett$GENDER == 1 & AutoSett$AGE >= 60),]

AutoSett[which(AutoSett$GENDER == 1 & AutoSett$AGE >= 60),]
```
# Task 3: Data Cleanup Stage --#
```{r setup, include=FALSE}
```
### Important to make a backup copy before doing data manipulation
```{r setup, include=FALSE}
AutoSettBackup <- AutoSett
AutoSettBackup
```
#-- Change SETTNUM to a factor variable, not integer
```{r setup, include=FALSE}
str(AutoSett)
AutoSett$SETTNUM <- factor(AutoSett$SETTNUM)
str(AutoSett)
```
# Task 4: Fixing missing data with known information --#
```{r setup, include=FALSE}
```
###Return rows with any NAs
```{r setup, include=FALSE}
AutoSett[!complete.cases(AutoSett),] 
```
### Get rows where gender is NA
```{r setup, include=FALSE}
AutoSett[!complete.cases(AutoSett$GENDER),] #Return rows where gender = NA
```
### Replace the NAs for the gender column
```{r setup, include=FALSE}
AutoSett[AutoSett$SETTNUM == "21", "GENDER"] <- 2
AutoSett[AutoSett$SETTNUM == "21", "GENDER"]
head(AutoSett)
AutoSett[AutoSett$SETTNUM == "47", "GENDER"] <- 2
AutoSett[AutoSett$SETTNUM == "1128", "GENDER"] <- 1
AutoSett[AutoSett$SETTNUM == "2018", "GENDER"] <- 1
AutoSett[AutoSett$SETTNUM == "2756", "GENDER"] <- 2
```
### SETTNUM	      GENDER
### 21	          2
### 47	          2
### 1128	        1
### 2018	        1
### 2756	        2
```{r setup, include=FALSE}
```
### Check your results
```{r setup, include=FALSE}
```
AutoSett[(AutoSett$SETTNUM == "21" | AutoSett$SETTNUM == "47" | 
            AutoSett$SETTNUM == "1128" | AutoSett$SETTNUM == "2018" |
           AutoSett$SETTNUM == "2756"),]
```{r setup, include=FALSE}
```
# Task 5: Fix SEATBELT missing data using median impute method
```{r setup, include=FALSE}
AutoSett[!complete.cases(AutoSett),] 
summary(AutoSett)

median(AutoSett$SEATBELT, na.rm = TRUE)

Seatbelt.Median <- median(AutoSett$SEATBELT, na.rm = TRUE)

AutoSett[is.na(AutoSett$SEATBELT),]
#37, 69, 120

AutoSett[AutoSett$SETTNUM == "37", "SEATBELT"] <- Seatbelt.Median

AutoSett[is.na(AutoSett$SEATBELT),"SEATBELT"] <- Seatbelt.Median
```
## Check your results
```{r setup, include=FALSE}
```
AutoSett[(AutoSett$SETTNUM == "37" | AutoSett$SETTNUM == "69" |
            AutoSett$SETTNUM == "120"),]
```{r setup, include=FALSE}
```
## Check your results
```{r setup, include=FALSE}
AutoSett[!complete.cases(AutoSett),]
```
#  Task 6: Fix missing data for AGE using average age by occupation 

### SETTNUM == 25 (engineering), 43 (Heath), 92 (Construction)

### example for SETTNUM == 25 who is an engineer
```{r setup, include=FALSE}
Engineers <- AutoSett[AutoSett$Occupation == "Engineering",]
Engineers
mean(Engineers$AGE)
mean(Engineers$AGE,na.rm = TRUE)
Engineer.Mean <- mean(Engineers$AGE,na.rm = TRUE)
Engineer.Mean

AutoSett[!complete.cases(AutoSett),]
```
###SETTNUM == 25
```{r setup, include=FALSE}
AutoSett[AutoSett$Occupation == "Engineering" & is.na(AutoSett$AGE),
         "AGE"] <- Engineer.Mean

AutoSett[AutoSett$SETTNUM == "25",]
```

```{r setup, include=FALSE}
AutoSett[!complete.cases(AutoSett),]
```
### Health: SETTNUM == 43
```{r setup, include=FALSE}
Health <- AutoSett[AutoSett$Occupation == "Health",]
Health

HealthAge.Mean <- mean(Health$AGE, na.rm = TRUE)
HealthAge.Mean

AutoSett[AutoSett$Occupation == "Health" & is.na(AutoSett$AGE), "AGE"] <- HealthAge.Mean

AutoSett[!complete.cases(AutoSett),]

AutoSett[!complete.cases(AutoSett),]

AutoSett[AutoSett$SETTNUM == "92",]
```
### Construction: SETTNUM == 92
```{r setup, include=FALSE}
Construction <- AutoSett[AutoSett$Occupation == "Construction",]
Construction

ConstructionAge.Mean <- mean(Construction$AGE, na.rm = TRUE)
ConstructionAge.Mean

AutoSett[AutoSett$Occupation == "Construction" & is.na(AutoSett$AGE),
         "AGE"] <- ConstructionAge.Mean
```
# Task 7: Clean messy data in SETTLEMENT column by replacement --#
```{r setup, include=FALSE}
head(AutoSett, n=20)
```
### Fix SETTLEMENT data for SETTNUM == 10 ($,), 37(dollars), 71 (-)
```{r setup, include=FALSE}
```
### gsub() does a replacement for all matches
```{r setup, include=FALSE}
AutoSett[AutoSett$SETTNUM == "10",] 
AutoSett$SETTLEMENT <- gsub(",","",AutoSett$SETTLEMENT)
```
### head(AutoSett)
```{r setup, include=FALSE}
AutoSett$SETTLEMENT <- gsub("\\$","",AutoSett$SETTLEMENT)

AutoSett[AutoSett$SETTNUM == "10","SETTLEMENT"] <- 42586
```
### Now fix the SETTLEMENT entries for SETTNUM == 37 and 71
### For SETTNUM == 37, remove the " dollars" from the value
### For SETTNUM == 71, change the sign from negative to positive

####For SETTNUM == 37
```{r setup, include=FALSE}
AutoSett[AutoSett$SETTNUM == "37",] 
AutoSett$SETTLEMENT <- gsub(" dollars","",AutoSett$SETTLEMENT)
```
###For SETTNUM == 71
```{r setup, include=FALSE}
AutoSett[AutoSett$SETTNUM == "71",]

AutoSett[AutoSett$SETTNUM == "71","SETTLEMENT"] <- 37964


knitr::opts_chunk$set(echo = TRUE)
```
